---
title: "Thesis"
author: "Sumit Anand"
date: "2024-04-17"
output: html_document
---

```{r }
library(pacman)
pacman::p_load(tidyverse,survival,haven,survminer, ggplot2, broom, knitr, gtsummary,descr, table1, kableExtra, dplyr, BiocManager,readxl, Hmisc, gridExtra, MASS, lme4, sandwich, merDeriv, ggdag, tidygraph, ggraph,gtsummary, gt, magrittr,htmlwidgets, tidyr,scales, webshot, survival, stargazer, sjPlot, effects, ggeffects)
```

```{r}
?MASS

DATA1<- read_excel("C:/Users/sumit/OneDrive/Desktop/thesis/code/Pre_survey.xlsx")
is.data.frame(DATA1)

DATA2<- read_excel("C:/Users/sumit/OneDrive/Desktop/thesis/code/post_survey1.xlsx")
is.data.frame(DATA1)



# Save original column names for reference
original_names <- colnames(DATA1)

# Define new short names for columns in DATA1
short_colnames <- c(  "SchoolCity",  "SchoolID", "UniqueID", "Age", "LivingArea", "Grade", "EduGoal",
  "OnlineSpend8",  "OnlineSpend9",  "Books10",  "Books11",  "HomeItems12",  "HomeItems13", "FamilyCar14", "FamilyCar15", "PeopleHome", "RoomsHome", "GuardianEdu", "HomeTech19", "HomeTech20",
  "TechSkill",  "AIExp22",  "AIExp23",  "AIContext",  "TechUse",  "TechComfort26", "TechComfort27",
  "InternetType", "JobKnow29", "JobKnow30", "CareerGuide31", "CareerGuide32", "CareerConf33",
  "CareerConf34", "JobImportance35", "JobImportance36", "JobPlan37", "JobPlan38", "LearnPref",
  "JobSearch40",  "JobSearch41",  "SchoolHelp42",  "SchoolHelp43",  "Treatment"
)


# Rename the dataframe columns
colnames(DATA1) <- short_colnames
colnames(DATA1) <- paste0("pre_", colnames(DATA1))


# Define new short names for the columns
short_colnames2 <- c(  "ID",  "SchoolCity",  "SchoolID",
  "UniqueID",  "DesiredEdu",  "TechSkill",  "TalkedToAI_7",
  "TalkedToAI_8",  "ComfortWithTech_9",  "ComfortWithTech_10",
  "ContextUsedAI",  "ChatGPTUseFreq_12",  "ChatGPTUseFreq_13",
  "ChatGPTFeatures","FutureCareerCertainty_15","FutureCareerCertainty_16",  "CareerKnowledge_17",  "CareerKnowledge_18",  "JobPlanConfidence_19",  "JobPlanConfidence_20",  "ChatGPTImportance_21",  "ChatGPTImportance_22",  "ChatGPTImpact_23", "ChatGPTImpact_24"
)

# Apply these new names to your dataframe
colnames(DATA2) <- short_colnames2
#subset data 1


DATA1 <- DATA1 %>%
  mutate(Outcome0 = case_when(
   pre_EduGoal == "Bachelor's degree" ~ "4- Undergraduate",
    pre_EduGoal == "High school diploma or equivalent" ~ "2- High School",
   pre_EduGoal == "Less than high school" ~ "1- Less than High School",
    pre_EduGoal == "Master's degree or higher" ~ " 5- Postgraduate",
    pre_EduGoal == "Not sure" ~ "0 - Uncertain",
   pre_EduGoal == "Vocational or technical training" ~ "3- TVET"
  )) %>% 
  mutate(Outcome0 = as.factor(Outcome0))


DATA2 <- DATA2 %>%
  mutate(Outcome1 = case_when(
    DesiredEdu == "Bachelor's degree" ~ "4- Undergraduate",
    DesiredEdu == "High school diploma or equivalent" ~ "2- High School",
    DesiredEdu == "Less than high school" ~ "1- Less than High School",
    DesiredEdu == "Master's degree or higher" ~ " 5- Postgraduate",
    DesiredEdu == "Not sure" ~ "0 - Uncertain",
    DesiredEdu == "Vocational or technical training" ~ "3 - TVET"
  )) %>% 
  mutate(Outcome1 = as.factor(Outcome1))



table(is.na(DATA1$pre_SchoolHelp43))

```


```{r}
# Recode GuardianEdu into 5 categories
DATA1 <- DATA1 %>%
  mutate(ParentsEdu = case_when(
    pre_GuardianEdu %in% c("Not sure / Prefer not to say") ~ "1-Not sure",
    pre_GuardianEdu %in% c("Less than high school") ~ "2-Less than high school",
    pre_GuardianEdu %in% c("High school diploma or equivalent") ~ "3-High school",
    pre_GuardianEdu %in% c("Vocational or technical training", "Bachelor's degree", "Master's degree or higher") ~ "4-Tertiary education",
    TRUE ~ "Other" # Catch-all for any unlisted categories
  ))

# Convert the new ParentsEdu variable to a factor with ordered levels
DATA1$ParentsEdu <- factor(DATA1$ParentsEdu, levels = c(
  "1-Not sure",
  "2-Less than high school",
  "3-High school",
  "4-Tertiary education"
))

DATA1$NewAge <- cut(DATA1$pre_Age, 
                    breaks = c(12, 13, 15, Inf), 
                    labels = c("12", "13-14", "15+"), 
                    right = FALSE, include.lowest = TRUE)

#AI experience 

DATA1$pre_AIExp23[DATA1$pre_AIExp23 == "Maybe"] <- "No"

table(DATA1$pre_AIExp23)

#Internet Type at Home 
table(DATA1$pre_InternetType)

# Replace 'Mobile data' and 'Wifi at home' with 'Yes'
DATA1$pre_InternetType <- replace(DATA1$pre_InternetType, DATA1$pre_InternetType %in% c("Mobile data", "Wifi at home"), "Yes")

# Replace 'None' with 'No'
DATA1$pre_InternetType <- replace(DATA1$pre_InternetType, DATA1$pre_InternetType == "None", "No")


sum(is.na(DATA1$pre_CareerGuide32))

DATA4 <- DATA1 %>%
  filter(!is.na(pre_Books10) & !is.na(pre_HomeItems12) & !is.na(pre_FamilyCar14) & !is.na(pre_PeopleHome) & !is.na(pre_RoomsHome)) 



# Define weights
weights <- c(pre_Books10 = 0.25, pre_HomeItems12 = 0.25, pre_FamilyCar14 = 0.25, pre_PeopleHome = 0.25)

# Normalize (scale) each variable
DATA4$NormalizedBooks10 <- DATA4$pre_Books10 / max(DATA4$pre_Books10, na.rm = TRUE) * weights['pre_Books10']
DATA4$NormalizedHomeItems12 <- DATA4$pre_HomeItems12 / max(DATA4$pre_HomeItems12, na.rm = TRUE) * weights['pre_HomeItems12']
DATA4$NormalizedFamilyCar14 <- DATA4$pre_FamilyCar14 / max(DATA4$pre_FamilyCar14, na.rm = TRUE) * weights['pre_FamilyCar14']

# Special treatment for PeopleHome as per your instruction
DATA4$AdjustedPeopleHome <- DATA4$pre_PeopleHome / DATA4$pre_RoomsHome
DATA4$NormalizedPeopleHome <- DATA4$AdjustedPeopleHome / max(DATA4$AdjustedPeopleHome, na.rm = TRUE) * weights['pre_PeopleHome']

# Sum the weighted components
DATA4$Income <- DATA4$NormalizedBooks10 + DATA4$NormalizedHomeItems12 + 
                DATA4$NormalizedFamilyCar14 + DATA4$NormalizedPeopleHome
hist(DATA4$Income, main = "Histogram of Income", xlab = "Income", col = "lightblue")

# Apply a log transformation to Income to reduce skewness
# Adding 1 to avoid log(0) which is undefined
DATA4$LogIncome <- log(DATA4$Income + 1)

# Round the transformed Income to two decimal places for cleaner display
DATA4$LogIncome <- round(DATA4$LogIncome, 2)

DATA4 <- DATA4 %>%
  filter(!is.na(LogIncome))

#Joining the two datasets

DATA3 <- left_join(DATA4, DATA2, by = c("pre_UniqueID"= "UniqueID"))
table(DATA3$Outcome1)
DATA5 <- left_join(DATA1, DATA2, by = c("pre_UniqueID"= "UniqueID"))
table(DATA3$Outcome1)

DATA3<- DATA3 %>%
  filter(!is.na(Outcome1), !is.na(Outcome0), !is.na(pre_Treatment))
table (DATA3$LogIncome)

DATA5<- DATA5 %>%
  filter(!is.na(Outcome1), !is.na(Outcome0), !is.na(pre_Treatment))
table (DATA3$LogIncome)

#Ttest for continuous variables

t_test_age <- t.test(pre_Age ~ pre_Treatment, data = DATA1)
print(t_test_age)


#chi-square for categorical variables

DATA1$pre_GuardianEdu <- as.factor(DATA1$pre_GuardianEdu)
DATA1$pre_HomeTech19 <- as.factor(DATA1$pre_HomeTech19)

# Chi-squared test
chi_sq_test <- chisq.test(table(DATA1$pre_Treatment, DATA1$pre_HomeTech19))
print(chi_sq_test)


#Ttest for income

t_test_income <- t.test(LogIncome ~ pre_Treatment, data = DATA4)
print(t_test_income)
```


```{r}
# Define the data for the table with lines for clarity
participant_flow_text <- c(
  "Table 1: Participant Flow at Different Stages",
  "------------------------------------------------",
  "",
  "Enrollment",
  "Estimate of teachers who opted in: n = 40",
  "Participated: n = 18 (didn't roll in n = 22)",
  "Estimate of students at the time of enrollment: n = 769",
  "Participated: n = 751 (duplicates removed, n = 18; all students are between the ages of 13-18 from low-income private or government schools)",
  "------------------------------------------------",
  "",
  "Randomisation",
  "Allocated to the intervention group: n = 426",
  "Allocated to the control group: n = 325",
  "------------------------------------------------",
  "",
  "Follow-up and Analyses",
  "Post-intervention follow-up returned: n = 380 (duplicates removed n = 11 and non-matching IDs to pre-survey were removed = 26, final n = 343)",
  "Intervention group: n = 177 (response rate 41.5%)",
  "Control group: n = 166 (response rate 51%)",
  "Lost to follow up: n = 408",
  "------------------------------------------------"
)

# Print the output to console
cat(participant_flow_text, sep = "\n")

# Optionally, write the output to a text file
writeLines(participant_flow_text, "Participant_Flow.txt")


```


```{r}
selected_data <- DATA1 %>%
  select(ParentsEdu, NewAge, pre_Treatment, pre_EduGoal, pre_JobKnow30, pre_CareerConf34,  pre_AIExp23, 
          pre_TechComfort27, pre_SchoolHelp43, pre_CareerGuide32, pre_InternetType) %>%
  rename( "Age" = NewAge,
    "Parents' Education" = ParentsEdu,
    "Educational Goal" = pre_EduGoal,
    "Job Knowledge" = pre_JobKnow30,
    "Career Confidence" = pre_CareerConf34,
    "Experience with GenAI" = pre_AIExp23,
    "Internet Access" = pre_InternetType,
    "Tech Comfort" = pre_TechComfort27,
    "Is school Helping?" = pre_SchoolHelp43,
    "Career Guidance" = pre_CareerGuide32
    )


# Create the table using the selected data
table <- selected_data %>%
  tbl_summary(
    by = pre_Treatment, # assuming pre_Treatment is a variable in DATA1
    type = list(all_categorical() ~ "categorical"),
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"
                     
    )
  )

# Assume 'table' is your tbl_summary object
table_with_p <- table %>%
  add_p(test = all_categorical() ~ "chisq.test")  # Chi-squared test for categorical variables

# Convert the updated table with p-values to a gt object and add title/footnote
gt_table_with_p <- as_gt(table_with_p) %>%
  gt::tab_header(
    title = "Table 2: Descriptive Statistics by Treatment Group with P-values"
  ) %>%
  gt::tab_footnote(
    footnote = "Data from the pre-survey. P-values are based on Chi-squared tests."
  )

# Print the gt table with p-values
print(gt_table_with_p)

# Save your gt_table as an HTML file first
gt::gtsave(gt_table_with_p, "table.html")

# Use the webshot package to take a screenshot of the HTML file and save it as an image
webshot("table.html", file = "table.png", delay = 0.2)


```



```{r}

# Load necessary libraries
library(stargazer)

pre_AIExp23 <- as.factor(DATA1$pre_AIExp23)
pre_TechComfort26 <- as.factor(DATA1$pre_TechComfort26)
pre_InternetType <- as.factor(DATA1$pre_InternetType)

# Define your models
model <- glm(pre_AIExp23 ~ LogIncome, family = binomial, data = DATA4)

modela <- glm(pre_AIExp23 ~ LogIncome + pre_TechComfort26 + pre_InternetType, family = binomial, data = DATA4)

#print both model with stargazer
model_AccessAI <- stargazer(model, modela, type = "text", title = "Logistic Regression Models", align = TRUE, single.row = TRUE, header = FALSE, digits = 2, exponentiate = TRUE)

#save the output as text
writeLines(model_AccessAI, "model_output.txt")



```


```{r}
# Assuming your dataset is named data, 
# with the binary outcome variable named access_AI, 
# the binary independent variable named internet_access, 
# and the continuous independent variable named income.


DATA4$pre_AIExp23 <- as.factor(DATA4$pre_AIExp23)

#descriptive graph of AI experience with parents income 
AI_income <-ggplot(DATA4, aes(x = LogIncome, fill = pre_AIExp23)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Density Plot of AI Experience by Log Income",
    subtitle = "Data from pre survey (N = 529)",
    x = "Log Income",
    y = "Density of AI Experience",
    fill = "AI Experience"
  ) +
  theme_minimal()

# Save the combined plot as a PNG file
png("AI_income.png",  width = 1000, height = 450)
print(AI_income) # Ensure that the plot is printed to the PNG device
dev.off()

```


```{r}
# Logistic regression with income as the predictor for access to AI
model <- glm(pre_AIExp23 ~ LogIncome , family = binomial, data = DATA4)
# get the summary of the model
summary_model <- summary(model)

# get the coefficients, p-values and confidence intervals
coefficients <- summary_model$coefficients

# calculate odds ratios
odds_ratios <- exp(coefficients[, "Estimate"])

# calculate lower and upper confidence intervals for odds ratios
ci_lower <- exp(coefficients[, "Estimate"] - 1.96 * coefficients[, "Std. Error"])
ci_upper <- exp(coefficients[, "Estimate"] + 1.96 * coefficients[, "Std. Error"])

# get p-values
p_values <- coefficients[, "Pr(>|z|)"]

# create a data frame to nicely print the results
results <- data.frame(Variable = rownames(coefficients), 
                      OddsRatio = odds_ratios, 
                      LowerCI = ci_lower, 
                      UpperCI = ci_upper, 
                      PValue = p_values)

# print the results
print(results)



# Rename the variable in the results data frame
results$Variable <- factor(results$Variable,
                           levels = results$Variable[order(results$OddsRatio)],
                           labels = c("Intercept", "Log Income"))
                          
                  

# Create the plot
plot <- ggplot(results, aes(x = Variable, y = OddsRatio)) +
  geom_point(color = "darkred") +
  geom_errorbar(aes(ymin = LowerCI, ymax = UpperCI), width = 0.2) +
  coord_trans(y = "log10") +
  labs(
    title = " Odds Ratios with Confidence Intervals of Income and its influence on AI experience ", # New title of the figure
    subtitle = "Data is from pre survey (N = 529)", # New subtitle of the figure
    x = "Variables in the regression", # New x-axis label
    y = "Odds Ratio (log scale)" # y-axis label remains the same
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), # Center the plot title
        plot.subtitle = element_text(hjust = 0.5, face = "italic"), # Center the plot subtitle and make it italic
        axis.title.x = element_text(size = 12, face = "bold"), # Increase the size and boldness of x-axis label
        axis.title.y = element_text(size = 12, face = "bold")) # Increase the size and boldness of y-axis label

# Print the plot
print(plot)

# Create the plot
plot <- ggplot(results, aes(x = Variable, y = OddsRatio)) +
  geom_point(color = "darkred") +
  geom_errorbar(aes(ymin = LowerCI, ymax = UpperCI), width = 0.2) +
  coord_trans(y = "log10") +
  labs(
    title = " Odds Ratios with Confidence Intervals of Income and its influence on AI experience ", # New title of the figure
    subtitle = "Data is from pre survey (N = 529)", # New subtitle of the figure
    x = "Variables in the regression", # New x-axis label
    y = "Odds Ratio (log scale)" # y-axis label remains the same
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), # Center the plot title
        plot.subtitle = element_text(hjust = 0.5, face = "italic"), # Center the plot subtitle and make it italic
        axis.title.x = element_text(size = 12, face = "bold"), # Increase the size and boldness of x-axis label
        axis.title.y = element_text(size = 12, face = "bold")) # Increase the size and boldness of y-axis label



# Save the combined plot as a PNG file
png("plot1.png",  width = 1000, height = 450)
print(plot) # Ensure that the plot is printed to the PNG device
dev.off()


```


```{r}


modela <- glm(pre_AIExp23 ~ LogIncome + pre_TechComfort26 + pre_InternetType, family = binomial, data = DATA4)
summary_modela <- summary(modela)
# get the coefficients, p-values and confidence intervals
coefficients <- summary_modela$coefficients

# calculate odds ratios
odds_ratios <- exp(coefficients[, "Estimate"])

# calculate lower and upper confidence intervals for odds ratios
ci_lower <- exp(coefficients[, "Estimate"] - 1.96 * coefficients[, "Std. Error"])
ci_upper <- exp(coefficients[, "Estimate"] + 1.96 * coefficients[, "Std. Error"])

# get p-values
p_values <- coefficients[, "Pr(>|z|)"]

# create a data frame to nicely print the results
results <- data.frame(Variable = rownames(coefficients), 
                      OddsRatio = odds_ratios, 
                      LowerCI = ci_lower, 
                      UpperCI = ci_upper, 
                      PValue = p_values)

# print the results
print(results)

# Create gtsummary table
tbl <- tbl_regression(modela, exponentiate = TRUE) %>%
  modify_header(
    label ~ "Variable",
    estimate ~ "Estimate (Odds Ratio)",
    ci ~ "95% CI",
    p.value ~ "P-value"
  ) %>%
  modify_caption("Summary of GLM Model Results") %>%
  bold_labels()

# Convert to kable format
kable_table <- as_kable(tbl)

# Customize with kableExtra
styled_table <- kable_table %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    position = "left"
  ) %>%
  row_spec(0, bold = TRUE, color = "Black")  # Example to make the header blue and bold

save_kable(styled_table, "model_results.html")

# Rename the variable in the results data frame
results$Variable <- factor(results$Variable,
                           levels = results$Variable[order(results$OddsRatio)],
                            labels = c("Intercept", "Tech Comfort", "Internet Access", "Log Income"))
                    
                  

# Create the plot
plot <- ggplot(results, aes(x = Variable, y = OddsRatio)) +
  geom_point(color = "darkred") +
  geom_errorbar(aes(ymin = LowerCI, ymax = UpperCI), width = 0.2) +
  coord_trans(y = "log10") +
  labs(
    title = " Odds Ratios with Confidence Intervals of Income and its influence AI experience with Control", # New title of the figure
    subtitle = "Pre Survey Data Used (N= 529)", # New subtitle of the figure
    x = "Variables in the regression", # New x-axis label
    y = "Odds Ratio (log scale)" # y-axis label remains the same
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), # Center the plot title
        plot.subtitle = element_text(hjust = 0.5, face = "italic"), # Center the plot subtitle and make it italic
        axis.title.x = element_text(size = 12, face = "bold"), # Increase the size and boldness of x-axis label
        axis.title.y = element_text(size = 12, face = "bold")) # Increase the size and boldness of y-axis label

# Print the plot
print(plot)

# Save the combined plot as a PNG file
png("plot2.png",  width = 1000, height = 450)
print(plot) # Ensure that the plot is printed to the PNG device
dev.off()



```



```{r}


```


```{r}
DATA1$ParentsEdu <- factor(DATA1$ParentsEdu, levels = c("1-Not sure", "2-Less than high school", "3-High school", "4-Tertiary education", "Other"))
# Create a bar plot for EduGoals across ParentsEdu
edu_plot <- ggplot(DATA1, aes(x = ParentsEdu, fill = pre_EduGoal)) +
  geom_bar(position = "dodge", width = 0.7, color = "black") + # Add black border
  scale_fill_brewer(palette = "Dark2") + # Use Dark2 color palette
  labs(
    title = "Educational Goals across Parent's Education",
    subtitle = "Data Source: Pre-survey Data",
    x = "Parent's Education",
    y = "Count",
    fill = "Educational Goals"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.x = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold"))

# Print the plot
print(edu_plot)

# Save the combined plot as a PNG file
png("plot3.png",  width = 1000, height = 450)
print(edu_plot) # Ensure that the plot is printed to the PNG device
dev.off()


```


```{r}
# Create a bar plot for AI Experience across Age
ai_plot <- ggplot(DATA1, aes(x = NewAge, fill = pre_AIExp23)) +
  geom_bar(position = "dodge", width = 0.5, color = "black") + # Add black border
  scale_fill_brewer(palette = "Dark2") + # Use Dark2 color palette
  labs(
    title = "AI Experience across Age Groups",
    subtitle = "Data Source: Pre-survey Data",
    x = "Age",
    y = "Count",
    fill = "AI Experience"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.x = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold"))

# Print the plot
print(ai_plot)


# Save the combined plot as a PNG file
png("plot4.png",  width = 1000, height = 450)
print(ai_plot) # Ensure that the plot is printed to the PNG device
dev.off()

```


```{r}
# Create a bar plot for Tech Comfort across Age 
tech_plot <- ggplot(DATA1, aes(x = NewAge, fill = pre_TechComfort27)) +
  geom_bar(position = "dodge", width = 0.5, color = "black") + # Add black border
  scale_fill_brewer(palette = "Dark2") + # Use Dark2 color palette
  labs(
    title = "Tech Comfort across Age Groups",
    subtitle = "Data Source: Pre-survey Data",
    x = "Age",
    y = "Count",
    fill = "Tech Comfort"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.x = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold"))

# Print the plot
print(tech_plot)
print(tech_plot)
# Save the combined plot as a PNG file
png("techplot.png",  width = 1000, height = 450)
print(school_plot) # Ensure that the plot is printed to the PNG device
dev.off()
```


```{r}
DATA6 <- DATA1 %>% 
  drop_na(pre_SchoolHelp43, pre_CareerGuide32)

# Create a bar plot for School Help across Career Guidance
school_plot <- ggplot(DATA6, aes(x = pre_CareerGuide32, fill = pre_SchoolHelp43)) +
  geom_bar(position = "dodge", width = 0.5, color = "black") + # Add black border
  scale_fill_brewer(palette = "Dark2") + # Use Dark2 color palette
  labs(
    title = "School Help across Career Guidance",
    subtitle = "Data Source: Pre-survey Data",
    x = "Career Guidance",
    y = "Count",
    fill = "School Help"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.x = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold"))


print(school_plot)
# Save the combined plot as a PNG file
png("plot5.png",  width = 1000, height = 450)
print(school_plot) # Ensure that the plot is printed to the PNG device
dev.off()


```



```


```{r}

#Main Model 1 (Career Goals)
# Assuming that you want the second level of Outcome0 to be the reference category
DATA5$Outcome0 <- relevel(as.factor(DATA5$Outcome0), ref = "0 - Uncertain")
DATA3$Outcome0 <- relevel(as.factor(DATA3$Outcome0), ref = "0 - Uncertain")

# Now, fit the model with the new reference level
model1 <- polr(as.factor(Outcome1) ~ pre_Treatment + Outcome0 , data = DATA5)
summary(model1)




DATA3$ParentsEdu <-as.factor(DATA3$ParentsEdu)
table(DATA3$ParentsEdu)


model12 <- polr(as.factor(Outcome1) ~ pre_Treatment + Outcome0 + ParentsEdu, data = DATA3)
summary(model12)



#Main Model 1 (Career Goals)
DATA3$pre_SchoolID <- as.factor(DATA3$pre_SchoolID)
levels(DATA5$Outcome0)
# Fit the model
model1_2 <- glmer(Outcome1 ~ pre_Treatment + Outcome0 + LogIncome + ParentsEdu + (1 | pre_SchoolID), 
                  data = DATA3, 
                  family = binomial("logit"))
                  

#join models with stargazer



model1_Edugoals <- stargazer(model1,model12, model1_2, type = "text", 
          title = "Regression Analysis: Model Summary for Education Goals",
          ci.levels = .95,  # 95% CI
          single.row = TRUE,
          header = FALSE,
          digits = 2,
          apply.coef = exp,  # Exponentiate the coefficients
          column.labels = c("Model 1", "Model 2", "Model 3"))  # Rename the variables

#save the output as text
writeLines(model1_Edugoals, "model1_edugoals.txt")






```

```{r}

model1 <- polr(as.factor(Outcome1) ~ pre_Treatment + Outcome0 , data = DATA5)


# For example, if 'pre_Treatment' is a binary treatment indicator, we can use both levels
new_data <- expand.grid(pre_Treatment = c(0, 1), 
                        Outcome0 = levels(DATA3$Outcome0))
                    
                  

# Generate predicted probabilities
predicted_probabilities <- predict(model1, new_data, type = "probs")

# Convert to a long format for ggplot2
predicted_probabilities_long <- cbind(new_data, as.data.frame(predicted_probabilities))
predicted_probabilities_long <- reshape2::melt(predicted_probabilities_long, id.vars = c("pre_Treatment", "Outcome0"))


# Calculate cumulative probabilities
cumulative_probabilities <- t(apply(predicted_probabilities, 1, cumsum))

# Convert to a long format for ggplot2
cumulative_probabilities_long <- cbind(new_data, as.data.frame(cumulative_probabilities))
cumulative_probabilities_long <- reshape2::melt(cumulative_probabilities_long, id.vars = c("pre_Treatment", "Outcome0"))


# Plot with ggplot2
model1plot2 <- ggplot(cumulative_probabilities_long, aes(x = pre_Treatment, y = value, color = Outcome0)) + 
  geom_line(aes(group = Outcome0), size = 1) + # Increase line size
  scale_color_brewer(palette = "Set1") + # Change line colors
  facet_wrap(~ variable, scales = "free") +
  labs(
    title = "Cumulative Predicted Probabilities with controls for Education Goals", # New title of the figure
    subtitle = "Data Source: Pre & Post-survey Data", # Subtitle of the figure (data source)
    x = "Treatment Group", # New x-axis label
    y = "Cumulative Predicted Probability", # New y-axis label
    color = "Outcome Level" # New legend title
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"), # Center the plot title and make it bold
    plot.subtitle = element_text(hjust = 0.5), # Center the plot subtitle
    axis.title.x = element_text(size = 12, face = "bold"), # Increase the size and boldness of x-axis label
    axis.title.y = element_text(size = 12, face = "bold"), # Increase the size and boldness of y-axis label
    legend.title = element_text(face = "bold") # Make the legend title bold
  )

model1plot2


```



```

```


# Store table of coefficients
ctable <- coef(summary_model1)

# Calculate and store p-values
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2

# Combine coefficients and p-values into a data frame
ctable <- cbind(ctable, "p value" = p)


# Default method gives profiled CIs
ci <- confint(model1)

# Calculate odds ratios
odds_ratios <- exp(coef(model1))

# Combine odds ratios and their confidence intervals
or_ci <- exp(cbind(OR = coef(model1), ci))

# Add p-values to the final output
or_ci <- cbind(or_ci, "p value" = p)


# Print the results
print(or_ci)


# For example, if 'pre_Treatment' is a binary treatment indicator, we can use both levels
new_data1 <- expand.grid(pre_Treatment = c(0, 1), 
                        Outcome0 = levels(DATA3$Outcome0), 
                        LogIncome = DATA3$LogIncome)



# Generate predicted probabilities
predicted_probabilities <- predict(model1, new_data, type = "probs")

# Convert to a long format for ggplot2
predicted_probabilities_long <- cbind(new_data, as.data.frame(predicted_probabilities))
predicted_probabilities_long <- reshape2::melt(predicted_probabilities_long, id.vars = c("pre_Treatment", "Outcome0"))

# Plot with ggplot2
model1plot <- ggplot(predicted_probabilities_long, aes(x = pre_Treatment, y = value, fill = Outcome0)) + 
  geom_bar(stat = "identity", position = "dodge", color = "black") + # Add outline color to bars
  scale_fill_brewer(palette = "Set1") + # Change fill colors
  facet_wrap(~ variable, scales = "fixed") +
  labs(
    title = "Predicted Probabilities by Treatment Group for Education Goals", # New title of the figure
    subtitle = "Data Source: Post-survey Data", # Subtitle of the figure (data source)
    x = "Treatment Group", # New x-axis label
    y = "Predicted Probability", # New y-axis label
    fill = "Outcome Level" # New legend title
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"), # Center the plot title and make it bold
    plot.subtitle = element_text(hjust = 0.5), # Center the plot subtitle
    axis.title.x = element_text(size = 12, face = "bold"), # Increase the size and boldness of x-axis label
    axis.title.y = element_text(size = 12, face = "bold"), # Increase the size and boldness of y-axis label
    legend.title = element_text(face = "bold") # Make the legend title bold
  )


# Save the combined plot as a PNG file
png("model1.png",  width = 1000, height = 450)
print(model1plot) # Ensure that the plot is printed to the PNG device
dev.off()

# Calculate cumulative probabilities
cumulative_probabilities <- t(apply(predicted_probabilities, 1, cumsum))

# Convert to a long format for ggplot2
cumulative_probabilities_long <- cbind(new_data, as.data.frame(cumulative_probabilities))
cumulative_probabilities_long <- reshape2::melt(cumulative_probabilities_long, id.vars = c("pre_Treatment", "Outcome0"))


# Plot with ggplot2
model1plot2 <- ggplot(cumulative_probabilities_long, aes(x = pre_Treatment, y = value, color = Outcome0)) + 
  geom_line(aes(group = Outcome0), size = 1) + # Increase line size
  scale_color_brewer(palette = "Set1") + # Change line colors
  facet_wrap(~ variable, scales = "free") +
  labs(
    title = "Cumulative Predicted Probabilities without controls for Education Goals", # New title of the figure
    subtitle = "Data Source: Post-survey Data", # Subtitle of the figure (data source)
    x = "Treatment Group", # New x-axis label
    y = "Cumulative Predicted Probability", # New y-axis label
    color = "Outcome Level" # New legend title
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"), # Center the plot title and make it bold
    plot.subtitle = element_text(hjust = 0.5), # Center the plot subtitle
    axis.title.x = element_text(size = 12, face = "bold"), # Increase the size and boldness of x-axis label
    axis.title.y = element_text(size = 12, face = "bold"), # Increase the size and boldness of y-axis label
    legend.title = element_text(face = "bold") # Make the legend title bold
  )


# Save the combined plot as a PNG file
png("plot7.png",  width = 1000, height = 450)
print(model1plot2) # Ensure that the plot is printed to the PNG device
dev.off()



```{r}
model1_3 <- glmer(Outcome1 ~ pre_Treatment*NewAge + Outcome0 + (1 | pre_SchoolID), 
                                    data = DATA5, 
                                    family = binomial("logit"))


summary(model1_3)



```


```{r}

DATA5$CareerKnowledge_18 <- as.factor(DATA5$CareerKnowledge_18)
DATA5$pre_JobKnow30 <- relevel(as.factor(DATA5$pre_JobKnow30), ref = "Nothing at all")
DATA3$CareerKnowledge_18 <- as.factor(DATA3$CareerKnowledge_18)
DATA3$pre_JobKnow30 <- relevel(as.factor(DATA3$pre_JobKnow30), ref = "Nothing at all")



#Normal Model
model2_1 <- polr(CareerKnowledge_18~pre_Treatment + pre_JobKnow30, data = DATA5)

#predicted probabilities
new_data2 <- expand.grid(pre_Treatment = c(0, 1), 
                        pre_JobKnow30 = levels(DATA3$pre_JobKnow30))
                        
predicted_probabilities2 <- predict(model2_1, new_data2, type = "probs")

# Convert to a long format for ggplot2
predicted_probabilities_long2 <- cbind(new_data2, as.data.frame(predicted_probabilities2))

predicted_probabilities_long2 <- reshape2::melt(predicted_probabilities_long2, id.vars = c("pre_Treatment", "pre_JobKnow30"))

#create culmulative probabilities
cumulative_probabilities2 <- t(apply(predicted_probabilities2, 1, cumsum))

# Convert to a long format for ggplot2
cumulative_probabilities_long2 <- cbind(new_data2, as.data.frame(cumulative_probabilities2))

cumulative_probabilities_long2 <- reshape2::melt(cumulative_probabilities_long2, id.vars = c("pre_Treatment", "pre_JobKnow30"))

# Plot with ggplot2 for cumulative probabilities

model2plot2 <- ggplot(cumulative_probabilities_long2, aes(x = pre_Treatment, y = value, color = pre_JobKnow30)) + 
  geom_line(aes(group = pre_JobKnow30), size = 1) +
  scale_color_discrete(labels = c("Nothing at all", "A little", "Some", "A lot")) +# Increase line size
  scale_color_brewer(palette = "Set1") + # Change line colors
  facet_wrap(~ variable, scales = "free") +
  labs(
    title = "Cumulative Predicted Probabilities without controls for Career Knowledge", # New title of the figure
    subtitle = "Data Source: Pre & Post-survey Data", # Subtitle of the figure (data source)
    x = "Treatment Group", # New x-axis label
    y = "Cumulative Predicted Probability", # New y-axis label
    color = "Job Knowledge" # New legend title
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"), # Center the plot title and make it bold
    plot.subtitle = element_text(hjust = 0.5), # Center the plot subtitle
    axis.title.x = element_text(size = 12, face = "bold"), # Increase the size and boldness of x-axis label
    axis.title.y = element_text(size = 12, face = "bold"), # Increase the size and boldness of y-axis label
    legend.title = element_text(face = "bold") # Make the legend title bold
  )
  
  model2plot2



#model with income
model22 <- polr(CareerKnowledge_18 ~ pre_Treatment + pre_JobKnow30 + LogIncome +ParentsEdu, data = DATA3)
summary(model22)

#model with fixed effects
model2_2 <- glmer(CareerKnowledge_18 ~ pre_Treatment + pre_JobKnow30 + LogIncome + ParentsEdu + (1 | pre_SchoolID), 
                data = DATA3, 
                family = binomial("logit"))

summary(model2_2)


#combine all three models in stargazer

table(DATA3$pre_JobKnow30)

model2_Career <- stargazer(model2_1,model22, model2_2, type = "text", 
          title = "Regression Analysis: Model Summary for Career Knowledge",
          ci.levels = .95,  # 95% CI
          single.row = TRUE,
          header = FALSE,
          digits = 2,
          apply.coef = exp,  # Exponentiate the coefficients
          column.labels = c("Model 1", "Model 2", "Model 3"))  # Rename the variables


#Model with interaction

model2_3 <- glmer(CareerKnowledge_18 ~ pre_Treatment + pre_JobKnow30 + NewAge + (1 | pre_SchoolID), 
                data = DATA5, 
                family = binomial("logit"))

summary(model2_3)




```


```{r}

DATA5$FutureCareerCertainty_16 <- as.factor(DATA5$FutureCareerCertainty_16)
DATA5$pre_CareerConf34 <- as.factor(DATA5$pre_CareerConf34)
DATA3$FutureCareerCertainty_16 <- as.factor(DATA3$FutureCareerCertainty_16)
DATA3$pre_CareerConf34 <- as.factor(DATA3$pre_CareerConf34)

model3_1 <- polr(FutureCareerCertainty_16 ~ pre_Treatment + pre_CareerConf34, data = DATA5)

#predicted probabilities
new_data3 <- expand.grid(pre_Treatment = c(0, 1), 
                        pre_CareerConf34 = levels(DATA3$pre_CareerConf34))
                        
predicted_probabilities3 <- predict(model3_1, new_data3, type = "probs")

# Convert to a long format for ggplot2
predicted_probabilities_long3 <- cbind(new_data3, as.data.frame(predicted_probabilities3))

predicted_probabilities_long3 <- reshape2::melt(predicted_probabilities_long3, id.vars = c("pre_Treatment", "pre_CareerConf34"))

#create culmulative probabilities

cumulative_probabilities3 <- t(apply(predicted_probabilities3, 1, cumsum))

# Convert to a long format for ggplot2

cumulative_probabilities_long3 <- cbind(new_data3, as.data.frame(cumulative_probabilities3))

cumulative_probabilities_long3 <- reshape2::melt(cumulative_probabilities_long3, id.vars = c("pre_Treatment", "pre_CareerConf34"))

# Plot with ggplot2 for cumulative probabilities

model3plot2 <- ggplot(cumulative_probabilities_long3, aes(x = pre_Treatment, y = value, color = pre_CareerConf34)) + 
  geom_line(aes(group = pre_CareerConf34), size = 1) +
  scale_color_discrete(labels = c("Not at all", "A little", "Some", "A lot")) +# Increase line size
  scale_color_brewer(palette = "Set1") + # Change line colors
  facet_wrap(~ variable, scales = "free") +
  labs(
    title = "Cumulative Predicted Probabilities without controls for Career Confidence", # New title of the figure
    subtitle = "Data Source: Post-survey Data", # Subtitle of the figure (data source)
    x = "Treatment Group", # New x-axis label
    y = "Cumulative Predicted Probability", # New y-axis label
    color = "Career Confidence" # New legend title
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"), # Center the plot title and make it bold
    plot.subtitle = element_text(hjust = 0.5), # Center the plot subtitle
    axis.title.x = element_text(size = 12, face = "bold"), # Increase the size and boldness of x-axis label
    axis.title.y = element_text(size = 12, face = "bold"), # Increase the size and boldness of y-axis label
    legend.title = element_text(face = "bold") # Make the legend title bold
  )
  
  model3plot2
  

model3_2 <- polr(FutureCareerCertainty_16 ~ pre_Treatment + pre_CareerConf34 + LogIncome + ParentsEdu, data = DATA3)

summary(model3_1)


model3_3 <- glmer(FutureCareerCertainty_16 ~ pre_Treatment + pre_CareerConf34 + LogIncome + ParentsEdu + (1 | pre_SchoolID), 
                data = DATA3, 
                family = binomial("logit"))

summary(model3_2)

#summary with stargazer
model3_Career <- stargazer(model3_1,model3_2, model3_3, type = "text", 
          title = "Regression Analysis: Model Summary for Career Confidence",
          ci.levels = .95,  # 95% CI
          single.row = TRUE,
          header = FALSE,
          digits = 2,
          apply.coef = exp,  # Exponentiate the coefficients
          column.labels = c("Model 1", "Model 2", "Model 3"))  # Rename the variables


#interaction model 
DATA5$NewAge <- as.factor(DATA5$NewAge)

model3_1in <- polr(FutureCareerCertainty_16 ~ pre_Treatment*NewAge + pre_CareerConf34, data = DATA5)
model2_1in <- polr(CareerKnowledge_18 ~ pre_Treatment*NewAge + pre_JobKnow30, data = DATA5)
model1in <- polr(as.factor(Outcome1) ~ pre_Treatment*NewAge + Outcome0 , data = DATA5)

#summary with stargazer
model3_Career <- stargazer(model3_1in,model2_1in, model1in, type = "text", 
          title = "Regression Analysis: Model Summary with Interaction",
          ci.levels = .95,  # 95% CI
          single.row = TRUE,
          header = FALSE,
          digits = 2,
          apply.coef = exp,  # Exponentiate the coefficients
          column.labels = c("Model 1", "Model 2", "Model 3"))  # Rename the variables




